---
title: "VCDB Analysis Notebook"
knit: (function(input_file, encoding) {
  out_dir <- 'docs'; 
  rmarkdown::render(input_file, 
  encoding=encoding, 
  output_file=file.path(dirname(input_file),
  out_dir, 'index.html'))
  })
output:
  html_notebook:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_depth: 3
    toc_float: yes
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# Setup

Load relevant libraries, set some notebook settings, and tweak default format for graphs.

```{r setup, eval=FALSE}
# libraries ----
library(tidyverse)
library(here)
library(verisr)

# code block options ----
knitr::opts_chunk$set(cache = FALSE)

# formatting ----
theme_set(theme_light())
```

Load data, specifically the validated incidents from https://github.com/vz-risk/VCDB.

```{r data}
# TODO download latest vcdb.dat
# TODO download latest data/validated files
# TODO download latest vcdb-merged.json

load(here::here("vcdb.dat"), verbose = TRUE)
vcdb_dir <- here::here("data", "validated")
schema_loc <- here::here("vcdb-merged.json")

vcdb <- verisr::json2veris(dir = vcdb_dir, schema = schema_loc)
```

# Summary

Incident counts since 2010, organized by `action` and `timeline.incident.year`.

```{r}
vcdb %>%
  getenumCI2020(
    enum = "action",
    by = "timeline.incident.year",
    ci.method = "bootstrap",
    na.rm = TRUE
  ) %>%
  filter(by >= 2010) %>%
  select(-method) %>%
  mutate(
    freq = scales::percent(freq, accuracy = 1),
    lower = scales::percent(lower, accuracy = 1),
    upper = scales::percent(upper, accuracy = 1)
  ) %>%
  DT::datatable()
```

When comparing to DBIR volumes, it's more appropriate to filter by `plus.dbir_year` instead of `timeline.incident.year`. For digging into breaches specifically, filter `attribute.confidentiality.data_disclosure.Yes`.

```{r}
vcdb %>% 
  filter(attribute.confidentiality.data_disclosure.Yes) %>% 
  getenumCI2020(
    enum = "action",
    by = "plus.dbir_year",
    ci.method = "bootstrap",
    na.rm = TRUE
  ) %>% 
  filter(by >= 2010) %>% 
  select(-method) %>% 
  mutate(freq = scales::percent(freq, accuracy = 1),
         lower = scales::percent(lower, accuracy = 1),
         upper = scales::percent(upper, accuracy = 1)) %>%
  DT::datatable()
```

There are a *lot* of variables, so exploring them can feel a bit... hack-y. Find all columns that include `disclosure`.

```{r}
grep("disclosure", names(vcdb), value = TRUE)
```

We can look at other high-level breakdowns, such as counts of breaches vs. incidents by DBIR publication years.

```{r}
vcdb %>%
  filter(plus.dbir_year >= 2014) %>% 
  group_by(attribute.confidentiality.data_disclosure.Yes) %>%
  count(plus.dbir_year) %>%
  ungroup() %>%
  rename(breach = attribute.confidentiality.data_disclosure.Yes) %>% 
mutate(breach = ifelse(breach, "Breach", "Incident")) %>%
  ggplot(mapping = aes(x = plus.dbir_year,
                       y = n,
                       group = breach)
         ) +
  geom_bar(mapping = aes(
    fill = breach
  ),
  stat = "identity"
  ) +
  geom_text(mapping = aes(
    y = n,
    label = n
  ),
  position = position_dodge(0.9),
  size = 3) +
  labs(title = "Breaches and Incidents by DBIR Year", 
       x = NULL, 
       y = NULL) +
  scale_x_continuous(
    breaks = seq(2014, 2021, by = 1)
  ) +
  scale_y_continuous(
    limits = c(0, 1000)
    ) +
  scale_fill_brewer() +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_blank()
  )
```

